---
title: "Machine Learning Course Project"
author: "Caroline Iurillo"
date: "November 19, 2015"
output: html_document
---
goal: The goal of this project is to predict the manner of performing unilateral dumbbell biceps curls based on data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants



Loading Data and Packages
```{r}
train<-read.csv("pml-training.csv", sep= ",", header=TRUE)
test<-read.csv("pml-testing.csv", sep= ",", header=TRUE)

library(caret)
set.seed(20)

```

subsetting data to remove columns with many NA's

```{r}
NAcount <- apply(train,2,function(x) {sum(is.na(x))}) #counts how many NA's are in each column
training <- train[,which(NAcount == 0)] # only include variables that have no NA's 

NAcount2 <- apply(test,2,function(x) {sum(is.na(x))}) 
testing<- test[,which(NAcount2 == 0)]

```

removing variables with near zero variance 

```{r}
nzv <- nearZeroVar(training,saveMetrics=TRUE)
training1 <- training[,nzv$nzv==FALSE]

nzv <- nearZeroVar(testing,saveMetrics=TRUE)
testing1 <- testing[,nzv$nzv==FALSE]

```


preping for random forest use by making sure data types are the same
```{r}
#remove user name - too many level, will mess with random forest
training2<-training1[c(-2)]#58 columns
testing2<-testing1[c(-2)]#58 columns
#remove time variable
training2<-training2[c(-4)]
testing2<-testing2[c(-4)]
#making the same type
table(sapply(training2[1,], class))
table(sapply(testing2[1,], class))
training2[-57] <- lapply(training2[-57], as.numeric)
training2$classe<-as.factor(training2$classe)
testing2[-51]<-lapply(testing2[-51],as.numeric)
```


Train and Test set for Cross Validation
```{r}
#cross validation, partitioned by classe
InTrain<-createDataPartition(y=training2$classe,p=.75, list=FALSE)
traintrain<-training2[InTrain,]
traintest<-training2[-InTrain,]

```

Random Forest model
used predict function for in sample error
From Confusion matrix you can tell the performance is very good.
```{r}

library(randomForest)
library(e1071)
rf<-randomForest(classe~., data=traintrain)
pred<-predict(rf,traintest)

confusionMatrix(pred,traintest$classe)

```